---
title: "PH Majlis Presiden MCO30 Statement"
author: "Frederik Paulus"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: 
  html_document: 
    keep_md: true
fig.width: 6
fig.asp: 0.618
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

# Load libraries
library(tidyverse)
library(ggplot2)
library(scales)
library(lubridate)
library(dplyr)
library(zoo)
library(hrbrthemes)
library(RColorBrewer)
library(modelr)

# Load REFSA logo
# refsa_logo <- png::readPNG("Logo.png")

# Load data

# ===== Google Mobility Reports =====
# https://www.google.com/covid19/mobility/
# https://www.apple.com/covid19/mobility (alternative source)
# Google provides data in two files (one per year), so merging them together here too. 
mobi_py <- read_csv("2020_MY_Region_Mobility_Report.csv")
mobi_cy <- read_csv("2021_MY_Region_Mobility_Report.csv")
mobi <- full_join(mobi_py, mobi_cy)

# Covid cases per million population
covid_cases_pm <- read_csv("https://raw.githubusercontent.com/owid/covid-19-data/master/public/data/jhu/new_cases_per_million.csv")

# ===== Covid test data (Our World in Data) =====
# https://github.com/owid/covid-19-data/blob/master/public/data/testing/covid-testing-all-observations.csv
covid_tests <- read_csv("https://raw.githubusercontent.com/owid/covid-19-data/master/public/data/testing/covid-testing-all-observations.csv")

```

## Daily Cases and Testing

Faced with the recent increase in the number of cases, the question is whether testing is sufficient. 

### Daily new cases

```{r covid-daily-new-cases, echo=FALSE}

# Basic graph of new cases per day for the last x months

# ===== Covid test and cases (Our World in Data) =====
# https://github.com/owid/covid-19-data/blob/master/public/data/jhu/new_cases_per_million.csv (Cases are per million population!)
# All Covid case data
covid_cases <- read_csv("https://raw.githubusercontent.com/owid/covid-19-data/master/public/data/jhu/new_cases.csv")

# Convert case data to a denormalised tibble. Pivot all columns except "Date" into a "country" column and a "cases" column. For PM data and all cases
covid_cases_p <- covid_cases %>% pivot_longer(!date, names_to = "country", values_to = "cases")

# Filter dataset for Malaysia only and for the last 6 months. Filter a bit longer for 7-day rolling average.
covid_cases_my <- filter(covid_cases_p, country == "Malaysia" & date > ymd("2020-10-15"))

# Add a 7-day average line to the data and filter again for the last 6 months
covid_cases_my <- covid_cases_my %>% mutate(cases_7ra = rollmean(cases, 7, fill = NA, align = "right")) %>% 
  filter(date > ymd("2020-10-31"))

# Plot plot plot
covid_cases_my_plot <- ggplot(data = covid_cases_my, aes(x = date)) +
  geom_col(aes(y = cases), color = "#FFDE00", fill = "#FFDE00") +
  geom_line(aes(y = cases_7ra), color = "#1252B3") +
  scale_x_date(name = "Date", date_breaks = "1 month", date_labels = "%b-%y") +
  scale_y_continuous(name = "Cases") +
  labs(title = "Daily new cases are ", subtitle = "Daily new cases and 7-day rolling average of new cases", caption = "Source: Our World in Data.")

covid_cases_my_plot


```

### Daily tests and test positivity rate

```{r covid-19-linear, echo=FALSE}

# ===== Data wrangling =====

# Convert case data to a denormalised tibble. Pivot all columns except "Date" into a "country" column and a "cases" column. For PM data and all cases
covid_cases_p <- covid_cases %>% pivot_longer(!date, names_to = "country", values_to = "cases")
covid_cases_pivot <- covid_cases_pm %>% pivot_longer(!date, names_to = "country", values_to = "cases_pm")

# Filter COVID-19 data for Malaysia and South Korea, from 2020-10-31 to 2021-02-28

# Define filter values
iso_codes <- c("MYS")
countries <- c("Malaysia")

date_from <- ymd("2021-01-01")
date_to <- ymd("2021-05-23")

# Define a tibble with the dates of MCO/CMCO restrictions
lockdowns <- tribble(
  ~lockdown, ~start, ~end,
  #--------|-------|-----
  "MCO-1", ymd("2020-03-18"), ymd("2020-05-04"),
  "CMCO-1", ymd("2020-05-04"), ymd("2020-06-09"),
  "CMCO-2", ymd("2020-10-14"), ymd("2020-12-05"),
  "MCO-2", ymd("2021-01-11"), ymd("2021-03-05"),
  "MCO-3", ymd("2021-05-25"), ymd("2021-06-16"),
)

# Filter case data and rename columns for ease of use; add an ISO code column for easy matching
covid_cases_m <- filter(covid_cases_pivot, country %in% countries & date > date_from & date < date_to)
covid_cases_m <- rename(covid_cases_m, 
                        record.date = `date`,
                        daily.cases.per.million = `cases_pm`)
covid_cases_m <- covid_cases_m %>% mutate(ISO.code = case_when(
  country == "Malaysia" ~ "MYS",
  TRUE ~ country
))

# Filter tests data according to parameters and only useful columns; rename columns for ease of use
covid_tests_m <- filter(covid_tests, covid_tests$`ISO code` %in% iso_codes & covid_tests$Date > date_from & covid_tests$Date < date_to) %>% select(c("ISO code", "Date", "Daily change in cumulative total", "Daily change in cumulative total per thousand", "7-day smoothed daily change", "7-day smoothed daily change per thousand", "Short-term positive rate", "Short-term tests per case")) 
covid_tests_m <- rename(covid_tests_m, 
      ISO.code = `ISO code`, 
      record.date = `Date`,
      daily.tests = `Daily change in cumulative total`, 
      daily.tests.per.thousand = `Daily change in cumulative total per thousand`, 
      smoothed.daily.change = `7-day smoothed daily change`, 
      smoothed.daily.change.per.thousand = `7-day smoothed daily change per thousand`, 
      st.pos.rate = `Short-term positive rate`, 
      st.tests.per.case = `Short-term tests per case`)

# Filter NA's and sort rows in ascending order (by Date)
covid_cases_m <- arrange(covid_cases_m, record.date)
covid_tests_m <- arrange(covid_tests_m, record.date)

# Merge cases and tests data
covid_all <- covid_tests_m %>% left_join(covid_cases_m, by = c("ISO.code", "record.date"))


```

### Evolution of testing 

```{r covid-daily-tests}

# New facet labels for countries
country.labs <- c("Malaysia")
names(country.labs) <- c("MYS")

# Plot daily tests (per thousand)
covid_dailytests_plot <- ggplot(data = covid_all, aes(x = record.date)) + 
  geom_bar(aes(y = daily.tests.per.thousand), stat = "identity") + 
  scale_x_date(name = "Date") +
  scale_y_continuous(name = "Count") + 
  labs(title = "Daily tests per thousand for South Korea and Malaysia", caption = "From 01-Oct-2020 to 07-May-2021. Source: Our World In Data.") +
  theme(legend.position = "none")

covid_dailytests_plot

```

### Test positivity rate

```{r covid-posrate}

# Plot test positivity rate
covid_posrate_plot <- ggplot(data = covid_all) + 
  geom_line(aes(x = record.date, y = st.pos.rate, color = ISO.code)) +
  scale_x_date(name = "Date") +
  scale_y_continuous(labels = scales::percent, name = "Percent") + 
  labs(title = "Short-term test positivity rate", caption = "From 01-Oct-2020 to 07-May-2021. Source: Our World In Data.") + 
  theme(legend.position = "none")

covid_posrate_plot + facet_wrap(~ ISO.code)

```

## Impact on the Economy

One way to gauge the impact on the economy is to use Google Mobility data to estimate the change in footfall in various locations where economic activity happens. 

```{r mobility, echo=FALSE}

# ===== Google Mobility Reports =====
# https://www.apple.com/covid19/mobility (alternative source)
# Google provides data in two files (one per year), so merging them together here too. 
mobi_py <- read_csv("2020_MY_Region_Mobility_Report.csv")
mobi_cy <- read_csv("2021_MY_Region_Mobility_Report.csv")
mobi <- full_join(mobi_py, mobi_cy)

# Filter the raw data to only include nationwide change and correct date range
mobi_nation <- filter(mobi, is.na(sub_region_1) & date > date_from & date < date_to)
mobi_nation <- rename(mobi_nation, 
                      retail.rec = "retail_and_recreation_percent_change_from_baseline", 
                      grocery.pharmacy = "grocery_and_pharmacy_percent_change_from_baseline", 
                      parks = "parks_percent_change_from_baseline", 
                      transit = "transit_stations_percent_change_from_baseline", 
                      workplaces = "workplaces_percent_change_from_baseline", 
                      residential = "residential_percent_change_from_baseline")

# New facet labels for place categories
dest_cats.labs <- c("Retail & Rec", "Grocery & Pharmacy", "Parks", "Transit Stations", "Workplaces", "Residential")
names(dest_cats.labs) <- c("retail.rec", "grocery.pharmacy", "parks", "transit", "workplaces", "residential")

# The data is shown in tabular format, not very useful for analysis, so need to pivot the different categories into one column
mobi_nation <- mobi_nation %>% pivot_longer(c("retail.rec", "grocery.pharmacy", "parks", "transit", "workplaces", "residential"), names_to = "category", values_to = "change.from.baseline")

```

### Mobility - Change from Baseline Across Categories

This chart shows the change in destinations relative to a pre-pandemic baseline, smoothed using a 7-day rolling average. Each time movement restrictions are announced, there is a noticeable drop in retail and recreation destinations as well as transit stations. 

```{r mobi-plot-all, echo=FALSE}

# Filter the table to pick up only national data; Add a column with 7-day rolling average
mobi_nation <- mobi_nation %>% arrange(desc(category)) %>%
  group_by(category) %>%
  mutate(change.from.baseline.smooth = rollmean(change.from.baseline, 7, fill = NA)) %>%
  ungroup()

# plot the data; re-using the labels defined in the mobility block. 
mobi_plot <- ggplot(data = mobi_nation) +
  geom_rect(data = filter(lockdowns, start > date_from), inherit.aes = FALSE, aes(xmin = start, xmax = end, ymin = -Inf, ymax = +Inf), fill= 'red', alpha = 0.1) +
  geom_area(aes(x = date, y = change.from.baseline.smooth, color = category, fill = category), alpha = 0.33) +
  scale_x_date(name = "Date", date_label = "%b-%y") +
  scale_y_continuous(name = "Change from baseline (%)") +
  labs(title = "Routing requests to most destinations remain below baseline", subtitle = "Change in routing requests compared to baseline (Jan-2020 to Feb-2020)", caption = "Shaded areas indicate MCO/CMCO restrictions.\nSource: Google Covid-19 Community Mobility Reports.") +  
  theme(legend.position = "none") +
  facet_wrap(~ category, labeller = labeller(category = dest_cats.labs)) +
  theme(panel.spacing.x = unit(4, "mm"))

mobi_plot 

```

### Do changes in mobility explain changes in GDP? 

Here we run a linear regression between the mobility data and the GDP, to look for a relation between the two. GDP per quarter is the dependent variable, while the change in mobility index from baseline per quarter is the independent variable. 

GDP data is only available quarterly, which limits the amount of data points. 

```{r regression}

# ===== Mobility data =====
# Bin all mobility observations in quarters, by date
mobi_nation_q <- mobi %>% filter(is.na(sub_region_1)) %>%
  mutate(quarter = case_when(
  (date >= ymd("2020-01-01") & date < ymd("2020-04-01")) ~ "1Q2020",
  (date >= ymd("2020-04-01") & date < ymd("2020-07-01")) ~ "2Q2020",
  (date >= ymd("2020-07-01") & date < ymd("2020-10-01")) ~ "3Q2020",
  (date >= ymd("2020-10-01") & date < ymd("2021-01-01")) ~ "4Q2020",
  (date >= ymd("2021-01-01") & date < ymd("2021-04-01")) ~ "1Q2021",  
  (date >= ymd("2021-04-01") & date < ymd("2021-07-01")) ~ "2Q2021"))  

# Rename the columns for easier work
mobi_nation_q <- rename(mobi_nation_q, 
                      retail.rec = "retail_and_recreation_percent_change_from_baseline", 
                      grocery.pharmacy = "grocery_and_pharmacy_percent_change_from_baseline", 
                      parks = "parks_percent_change_from_baseline", 
                      transit = "transit_stations_percent_change_from_baseline", 
                      workplaces = "workplaces_percent_change_from_baseline", 
                      residential = "residential_percent_change_from_baseline")

# The data is shown in tabular format, not very useful for analysis, so need to pivot the different categories into one column
mobi_nation_q <- mobi_nation_q %>% pivot_longer(c("retail.rec", "grocery.pharmacy", "parks", "transit", "workplaces", "residential"), names_to = "category", values_to = "change.from.baseline")

# Calculate an average value for economically relevant variables - exclude parks and residential
# Need to average across four categories and keep the other variables. Average over quarters.
mobi_nation_q <- mobi_nation_q %>% 
  filter(category %in% c("retail.rec", "grocery.pharmacy", "transit", "workplaces")) %>%
  group_by(country_region_code, country_region, quarter) %>%
  summarise(avg_delta = mean(change.from.baseline, na.rm = TRUE))

# ===== GDP Data =====
# Define a tibble with GDP data 
gdp_data <- tribble(
  ~quarter, ~gdp,
  #--|--
  "1Q2020", 344.7,
  "2Q2020", 289.7,
  "3Q2020", 351.6,
  "4Q2020", 357.9,
  "1Q2021", 343.1
)

# Join quarterly mobility data and quarterly GDP data
mobgdp_data <- left_join(mobi_nation_q, gdp_data, by = c("quarter" = "quarter"))

# Run a linear regression on GDP with the change in mobility
gdp_reg <- lm(gdp ~ avg_delta, data = mobgdp_data)

# Plot the actual data points and the model prediction
# First create a grid with data_grid() that contains all the unique variables. See here: https://r4ds.had.co.nz/model-basics.html#predictions

grid <- mobgdp_data %>% data_grid(avg_delta)

# then add the predictions
grid <- grid %>% add_predictions(gdp_reg)

# Last, plot everything
gdp_pred_plot <- ggplot(data = mobgdp_data, aes(x = avg_delta)) +
  geom_point(aes(y = gdp)) +
  geom_line(aes(y = pred), data = grid, colour = "red", size = 1)
  
gdp_pred_plot


```

## Comparing MCO 1 and MCO 3

On June 1st, the government announced a new MCO, purportedly as strict as the original MCO (MCO 1) in March 2020. It's worth comparing the state of the pandemic and the mobility in both cases, as well as reiterating the economic cost of the first, and the related support packages. 

```{r MCO-cases, echo=FALSE}

# Basic graph of new cases per day for the last x months

# Define a tibble with the dates of MCO/CMCO restrictions
lockdowns <- tribble(
  ~lockdown, ~start, ~end,
  #--------|-------|-----
  "MCO-1", ymd("2020-03-18"), ymd("2020-05-04"),
  "CMCO-1", ymd("2020-05-04"), ymd("2020-06-09"),
  "CMCO-2", ymd("2020-10-14"), ymd("2020-12-05"),
  "MCO-2", ymd("2021-01-11"), ymd("2021-03-05"),
  "MCO-3", ymd("2021-05-12"), ymd("2021-06-01"),
  "FMCO", ymd("2021-06-01"), ymd("2021-07-15")
)

# ===== Covid test and cases (Our World in Data) =====
# https://github.com/owid/covid-19-data/blob/master/public/data/jhu/new_cases_per_million.csv (Cases are per million population!)
# All Covid case data
daily_cases <- read_csv("https://raw.githubusercontent.com/owid/covid-19-data/master/public/data/jhu/new_cases.csv")

# Convert case data to a denormalised tibble. Pivot all columns except "Date" into a "country" column and a "cases" column. For PM data and all cases
daily_cases_p <- daily_cases %>% pivot_longer(!date, names_to = "country", values_to = "cases")

# Filter out Malaysia figures only
daily_cases_my <- filter(daily_cases_p, country == "Malaysia")

# Add a 7-day average line to the data and filter again for the last 6 months
daily_cases_my <- daily_cases_my %>% mutate(cases_7ra = rollmean(cases, 7, fill = NA, align = "right"))

# Create a vector of additional dates to model the end of the MCO.Take the planned end date + add 14 days
mco3_end <- ymd(lockdowns %>% filter(lockdown == "FMCO") %>% pull(end))
# figure out the max date in the period and fill in the gaps
mco3_gap <- interval(max(daily_cases_my$date), mco3_end)
mco3_gap_days <- mco3_gap/ddays(1) + 14
post_mco3_period <- max(daily_cases_my$date) + days(1:mco3_gap_days)
# Add it to the tibble of cases
daily_cases_my <- daily_cases_my %>% add_row(date = post_mco3_period)

# Add a column to the dataset to indicate whether it is MCO 1 or MCO 3
# Bin all daily case observations in MCO1 or MCO3, by date
daily_cases_my <- daily_cases_my %>% mutate(mco = case_when(
  (date >= ( ymd(lockdowns %>% filter(lockdown == "MCO-1") %>% pull(start)) - days(14) ) & date < ( ymd(lockdowns %>% filter(lockdown == "MCO-1") %>% pull(end)) + days(14) )) ~ "MCO1", 
  (date >= ( ymd(lockdowns %>% filter(lockdown == "MCO-3") %>% pull(start)) - days(14) ) & date < ( ymd(lockdowns %>% filter(lockdown == "FMCO") %>% pull(end)) + days(14) )) ~ "MCO3 / FMCO")) 

# Plot plot plot
daily_cases_my_plot <- ggplot(data = filter(daily_cases_my, mco != "NA"), aes(x = date)) +
  geom_col(aes(y = cases), color = "#FFDE00", fill = "#FFDE00") +
  geom_line(aes(y = cases_7ra), color = "#1252B3") +
  geom_vline(xintercept = ymd(lockdowns %>% filter(lockdown == "MCO-1") %>% pull(start)), color = "#999999") +
  geom_vline(xintercept = ymd(lockdowns %>% filter(lockdown == "MCO-1") %>% pull(end)), color = "#999999") +
  geom_vline(xintercept = ymd(lockdowns %>% filter(lockdown == "MCO-3") %>% pull(start)), color = "#999999") +
  geom_vline(xintercept = ymd(lockdowns %>% filter(lockdown == "FMCO") %>% pull(start)), color = "#999999") +
  geom_vline(xintercept = ymd(lockdowns %>% filter(lockdown == "FMCO") %>% pull(end)), color = "#999999") +
  scale_x_date(name = "Date", date_breaks = "7 days", date_labels = "%d-%b-%y") +
  scale_y_continuous(name = "Cases") +
  labs(title = "MCO 3 shows a plateau of new cases, but at a much higher base", subtitle = "Count and 7-day rolling average of daily new cases", caption = "MCO periods indicated by vertical lines.\nSource: Our World in Data.") +
  facet_wrap(~ mco, scales = "free_x") +
  theme(axis.text.x=element_text(angle=90, hjust=1))

daily_cases_my_plot
  

```



```{r MCO-daily-tests}

# New facet labels for countries
country.labs <- c("Malaysia")
names(country.labs) <- c("MYS")

# Filter tests data according to parameters and only useful columns; rename columns for ease of use
covid_mco_tests_m <- filter(covid_tests, covid_tests$`ISO code` %in% iso_codes) %>% select(c("ISO code", "Date", "Daily change in cumulative total", "Daily change in cumulative total per thousand", "7-day smoothed daily change", "7-day smoothed daily change per thousand", "Short-term positive rate", "Short-term tests per case")) 
covid_mco_tests_m <- rename(covid_mco_tests_m, 
      ISO.code = `ISO code`, 
      record.date = `Date`,
      daily.tests = `Daily change in cumulative total`, 
      daily.tests.per.thousand = `Daily change in cumulative total per thousand`, 
      smoothed.daily.change = `7-day smoothed daily change`, 
      smoothed.daily.change.per.thousand = `7-day smoothed daily change per thousand`, 
      st.pos.rate = `Short-term positive rate`, 
      st.tests.per.case = `Short-term tests per case`)

# Create a vector of additional dates to model the end of the MCO.Take the planned end date + add 14 days
mco3_end <- ymd(lockdowns %>% filter(lockdown == "FMCO") %>% pull(end))
# figure out the max date in the period and fill in the gaps
mco3_gap <- interval(max(covid_mco_tests_m$record.date), mco3_end)
mco3_gap_days <- mco3_gap/ddays(1) + 14
post_mco3_period <- max(covid_mco_tests_m$record.date) + days(1:mco3_gap_days)
# Add it to the tibble of cases
covid_mco_tests_m <- covid_mco_tests_m %>% add_row(record.date = post_mco3_period)

# Add a column to the dataset to indicate whether it is MCO 1 or MCO 3
# Bin all daily case observations in MCO1 or MCO3, by date
covid_mco_tests_m <- covid_mco_tests_m %>% mutate(mco = case_when(
  (record.date >= ( ymd(lockdowns %>% filter(lockdown == "MCO-1") %>% pull(start)) - days(14) ) & record.date < ( ymd(lockdowns %>% filter(lockdown == "MCO-1") %>% pull(end)) + days(14) )) ~ "MCO1", 
  (record.date >= ( ymd(lockdowns %>% filter(lockdown == "MCO-3") %>% pull(start)) - days(14) ) & record.date < ( ymd(lockdowns %>% filter(lockdown == "FMCO") %>% pull(end)) + days(14) ) ) ~ "MCO3 / FMCO")) 

# Plot daily tests (per thousand)
covid_mco_dailytests_plot <- ggplot(data = filter(covid_mco_tests_m, mco != "NA"), aes(x = record.date)) + 
  geom_vline(xintercept = ymd(lockdowns %>% filter(lockdown == "MCO-1") %>% pull(start)), color = "#999999") +
  geom_vline(xintercept = ymd(lockdowns %>% filter(lockdown == "MCO-1") %>% pull(end)), color = "#999999") +
  geom_vline(xintercept = ymd(lockdowns %>% filter(lockdown == "MCO-3") %>% pull(start)), color = "#999999") +
  geom_vline(xintercept = ymd(lockdowns %>% filter(lockdown == "FMCO") %>% pull(start)), color = "#999999") +
  geom_vline(xintercept = ymd(lockdowns %>% filter(lockdown == "FMCO") %>% pull(end)), color = "#999999") +
  geom_bar(aes(y = daily.tests.per.thousand), stat = "identity") + 
  scale_x_date(name = "Date", date_breaks = "7 days", date_labels = "%d-%b-%y") +
  scale_y_continuous(name = "Count") + 
  labs(title = "In proportion to the size of the outbreak, there is less testing in MCO 3 than MCO 1", subtitle = "Daily tests per thousand during MCO 1 and MCO 3", caption = "MCO periods indicated by vertical lines.\nSource: Our World In Data.") +
  theme(legend.position = "none") +
  theme(axis.text.x=element_text(angle=90, hjust=1))

covid_mco_dailytests_plot + facet_wrap(~ mco, scales = "free_x")

```


```{r}

# Plot daily tests (per thousand)
covid_mco_tpr_plot <- ggplot(data = filter(covid_mco_tests_m, mco != "NA"), aes(x = record.date)) + 
  geom_line(aes(y = st.pos.rate), color = "#1252B3") + 
  geom_vline(xintercept = ymd(lockdowns %>% filter(lockdown == "MCO-1") %>% pull(start)), color = "#999999") +
  geom_vline(xintercept = ymd(lockdowns %>% filter(lockdown == "MCO-1") %>% pull(end)), color = "#999999") +
  geom_vline(xintercept = ymd(lockdowns %>% filter(lockdown == "MCO-3") %>% pull(start)), color = "#999999") +
  geom_vline(xintercept = ymd(lockdowns %>% filter(lockdown == "FMCO") %>% pull(start)), color = "#999999") +
  geom_vline(xintercept = ymd(lockdowns %>% filter(lockdown == "FMCO") %>% pull(end)), color = "#999999") +
  scale_x_date(name = "Date", date_breaks = "7 days", date_labels = "%d-%b-%y") +
  scale_y_continuous(name = "Test positivity rate", labels = label_percent()) + 
  labs(title = "Test positivity rate quickly dropped in MCO 1 but not in MCO 3", subtitle = "Daily tests per thousand during MCO 1 and MCO 3", caption = "MCO periods indicated by vertical lines.\nSource: Our World In Data.") +
  theme(legend.position = "none") +
  theme(axis.text.x=element_text(angle=90, hjust=1)) +
  facet_wrap(~ mco, scales = "free_x")

covid_mco_tpr_plot


```

## Mobility during MCO 1 and MCO 3

Looking at Google changes in baseline. 

```{r mco-mobility, fig.asp = 1}

# Filter the raw data to only include nationwide change and correct date range
mobi_mco <- filter(mobi, is.na(sub_region_1))
mobi_mco <- rename(mobi_mco, 
                      retail.rec = "retail_and_recreation_percent_change_from_baseline", 
                      grocery.pharmacy = "grocery_and_pharmacy_percent_change_from_baseline", 
                      parks = "parks_percent_change_from_baseline", 
                      transit = "transit_stations_percent_change_from_baseline", 
                      workplaces = "workplaces_percent_change_from_baseline", 
                      residential = "residential_percent_change_from_baseline")

# New facet labels for place categories
dest_cats.labs <- c("Retail & Rec", "Grocery & Pharmacy", "Parks", "Transit Stations", "Workplaces", "Residential")
names(dest_cats.labs) <- c("retail.rec", "grocery.pharmacy", "parks", "transit", "workplaces", "residential")

# The data is shown in tabular format, not very useful for analysis, so need to pivot the different categories into one column
mobi_mco <- mobi_mco %>% pivot_longer(c("retail.rec", "grocery.pharmacy", "parks", "transit", "workplaces", "residential"), names_to = "category", values_to = "change.from.baseline")

# Filter the table to pick up only national data; Add a column with 7-day rolling average
mobi_mco <- mobi_mco %>% arrange(desc(category)) %>%
  group_by(category) %>%
  mutate(change.from.baseline.smooth = rollmean(change.from.baseline, 7, fill = NA, align = "right")) %>%
  ungroup()

# Add a column to the dataset to indicate whether it is MCO 1 or MCO 3
# Not doing the padding of -14d / +14d here, because the categories complicate things
# Bin all daily case observations in MCO1 or MCO3, by date
mobi_mco <- mobi_mco %>% mutate(mco = case_when(
  (date >= ( ymd(lockdowns %>% filter(lockdown == "MCO-1") %>% pull(start)) - days(14) ) & date < ( ymd(lockdowns %>% filter(lockdown == "MCO-1") %>% pull(end)) + days(14)) ) ~ "MCO1", 
  (date >= ( ymd(lockdowns %>% filter(lockdown == "MCO-3") %>% pull(start)) - days(14) ) & date < ( ymd(lockdowns %>% filter(lockdown == "FMCO") %>% pull(end)) + days(14) )) ~ "MCO3 / FMCO")) 

# plot the data; re-using the labels defined in the mobility block. 
mobi_plot <- ggplot() +
  geom_vline(xintercept = ymd(lockdowns %>% filter(lockdown == "MCO-1") %>% pull(start)), color = "#999999") +
  geom_vline(xintercept = ymd(lockdowns %>% filter(lockdown == "MCO-1") %>% pull(end)), color = "#999999") +
  geom_vline(xintercept = ymd(lockdowns %>% filter(lockdown == "MCO-3") %>% pull(start)), color = "#999999") +
  geom_vline(xintercept = ymd(lockdowns %>% filter(lockdown == "FMCO") %>% pull(start)), color = "#999999") +
  geom_vline(xintercept = ymd(lockdowns %>% filter(lockdown == "FMCO") %>% pull(end)), color = "#999999") +
  geom_area(data = filter(mobi_mco, mco != "NA" & category %in% c("retail.rec", "workplaces")), aes(x = date, y = change.from.baseline.smooth, color = category, fill = category), alpha = 0.33) +
  scale_x_date(name = "Date", date_label = "%d-%b-%y") +
  scale_y_continuous(name = "Change from baseline (%)") +
  labs(title = "Routing requests to most destinations are significantly higher in MCO 3", subtitle = "Change in routing requests compared to baseline (Jan-2020 to Feb-2020)", caption = "MCO periods indicated by vertical lines.\nSource: Google Covid-19 Community Mobility Reports.") +  
  theme(legend.position = "none") +
  facet_grid(rows = vars(category), cols = vars(mco), labeller = labeller(category = dest_cats.labs), scales = "free_x") +
  theme(panel.spacing.x = unit(4, "mm")) +
  theme(axis.text.x=element_text(angle=90, hjust=1))

mobi_plot 


```

