---
title: "REFSA COVID-19 Malaysia Response"
author: "REFSA Research Team"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: 
  html_document: 
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

# Load libraries
library(tidyverse)
library(ggplot2)
library(scales)
library(lubridate)
library(dplyr)

# Load REFSA logo
# refsa_logo <- png::readPNG("Logo.png")

# Load data
# ===== Google Mobility Reports =====
# https://www.apple.com/covid19/mobility
# mobi <- read.csv("https://covid19-static.cdn-apple.com/covid19-mobility-data/2007HotfixDev48/v2/en-us/applemobilitytrends-2020-05-04.csv", na.strings = "", fileEncoding = "UTF-8-BOM")
mobi <- read_csv("2020_MY_Region_Mobility_Report.csv")

# ===== Covid test and cases (Our World in Data) =====
# https://www.ecdc.europa.eu/en/publications-data/download-todays-data-geographic-distribution-covid-19-cases-worldwide
covid_data <- read_csv("covid-19-daily-tests-vs-daily-new-confirmed-cases-per-million.csv", col_types = cols(Entity = col_character(), Code = col_character(), Year = col_character(), Continent = col_character(), Date = col_date(), DailyTests = col_double(), DailyTestsNotes = col_character(), DailyCases = col_double()))

```


## Objective

Malaysia is currently struggling to deal with the third wave of the Covid-19 pandemic. This analysis aims to highlight a few data points that illuminate the response so far of the government, and where additional investment may be required. 

 1. Test positivity rate, and the need to perform more testing
 2. Changes in mobility, related to the positivity rate
 
## Test positivity rate

The test positivity rate is still quite elevated in Malaysia, above the minimum threshold recommended by the WHO (5%), and especially high compared to a successful country in managing Covid-19, South Korea. The following charts depict the evolution of number of new cases, number of tests, and positivity rate. All data is 7-day smoothed (rolling average); people tested are per 1,000 population; new daily cases are per million population.

Source: Our World In Data. 

```{r covid-19-linear, echo=FALSE}

# Filter COVID-19 data for Malaysia and South Korea
geo_ids <- c("MYS", "KOR")
covid_data_m <- filter(covid_data, Code %in% geo_ids)

# Only use data from September onwards
covid_data_m <- filter(covid_data_m, Date > ymd("2020-10-31") & Date < ymd("2021-01-23"))

# Calculate the test positivity rate
covid_data_MYd <- covid_data_m %>% mutate(DailyPosRate = (DailyCases / (DailyTests * 1000)) * 100)

# Filter NA's and sort rows in ascending order (by Date)
# covid_data_MYd <- filter(covid_data_MYd, ! is.na(DailyTests))
covid_data_MYd <- arrange(covid_data_MYd, Date)

# Calculate cumulative number of cases and deaths - not required here, plot is relative
# covid_data_MYd <- covid_data_MYd %>% group_by(geoId) %>% mutate(cumCases = cumsum(cases))
# covid_data_MYd <- covid_data_MYd %>% group_by(geoId) %>% mutate(cumDeaths = cumsum(deaths))

# covid_data_clean <- na.omit(covid_data_MYd)

```

```{r covid-plot, echo=FALSE}

# New facet labels for countries
country.labs <- c("South Korea", "Malaysia")
names(country.labs) <- c("KOR", "MYS")

# Plot MY COVID-19 cases
covid_dailytests_plot <- ggplot(data = covid_data_MYd, aes(x=Date)) + 
  geom_bar(aes(y = DailyCases / 10), stat = "identity", size = .1, fill = "#FFDE00", color = "#FFDE00", alpha = .6) +  
  geom_line(aes(y = DailyPosRate), color = "#1252B3") +
  scale_x_date(name = "Date") +
  scale_y_continuous(
    
    # Left axis
    name = "Test positivity rate (%)", breaks = seq(0,12,1),
    
    # Right axis
    sec.axis = sec_axis(~.*10, name = "Daily cases (per million)", breaks=seq(0,120,10))
  ) + 
  labs(title = "Test positivity rate and daily new cases", caption = "From 01-Nov-2020 to 22-Jan-2021. Daily new cases are 7-day rolling average. Source: Our World In Data.")
  
covid_dailytests_plot + facet_wrap(~ Code, labeller = labeller(Code = country.labs))

```

## Mobility
Analysis of mobility is based on Google's data for Malaysia. 
The charts below show the percent change compared to a baseline. The baseline is the median number of daily routing requests between early January to early February 2020. 

```{r mobility, echo=FALSE}

# Filter the raw data to only include nationwide change and correct date range
mobi_nation <- filter(mobi, is.na(sub_region_1) & date > ymd("2020-08-31"))

# The data is shown in tabular format, not very useful for analysis, so need to pivot the different categories into one column
mobi_nation_pivot <- mobi_nation %>% pivot_longer(c("retail_and_recreation_percent_change_from_baseline", "grocery_and_pharmacy_percent_change_from_baseline", "parks_percent_change_from_baseline", "transit_stations_percent_change_from_baseline", "workplaces_percent_change_from_baseline", "residential_percent_change_from_baseline"), names_to = "category", values_to = "change_from_baseline")

# plot the data
mobi_plot <- ggplot(data = mobi_nation_pivot) +
  geom_line(aes(x = date, y = change_from_baseline, color = category)) +
  scale_x_date(name = "Date") +
  theme(legend.position = "none")

mobi_plot + facet_wrap(~ category)

```
