---
title: "REFSA COVID-19 Response—MCO 2.0"
author: "REFSA Research Team"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: 
  html_document: 
    keep_md: true
fig.width: 6
fig.asp: 0.618
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

# Load libraries
library(tidyverse)
library(ggplot2)
library(scales)
library(lubridate)
library(dplyr)
library(zoo)
library(hrbrthemes)
library(RColorBrewer)

# Load REFSA logo
# refsa_logo <- png::readPNG("Logo.png")

# Load data

# ===== Google Mobility Reports =====
# https://www.apple.com/covid19/mobility (alternative source)
# Google provides data in two files (one per year), so merging them together here too. 
mobi_py <- read_csv("2020_MY_Region_Mobility_Report.csv")
mobi_cy <- read_csv("2021_MY_Region_Mobility_Report.csv")
mobi <- full_join(mobi_py, mobi_cy)

# ===== Covid test and cases (Our World in Data) =====
# https://github.com/owid/covid-19-data/blob/master/public/data/jhu/new_cases_per_million.csv (Cases are per million population!)
# All Covid case data
covid_cases <- read_csv("https://raw.githubusercontent.com/owid/covid-19-data/master/public/data/jhu/new_cases.csv")

# Covid cases per million population
covid_cases_pm <- read_csv("https://raw.githubusercontent.com/owid/covid-19-data/master/public/data/jhu/new_cases_per_million.csv")

# ===== Covid test data (Our World in Data) =====
# https://github.com/owid/covid-19-data/blob/master/public/data/testing/covid-testing-all-observations.csv
covid_tests <- read_csv("https://raw.githubusercontent.com/owid/covid-19-data/master/public/data/testing/covid-testing-all-observations.csv")

```


This tracker brings together a number of oft-referred to statistics to monitor the evolution of the Covid-19 pandemic in Malaysia, and highlight some interesting data points. As of now, the following data is included: 

 1. Daily count of tests, cases and test positivity rate
 2. Mobility data, based on Google's Covid-19 location data
 
## Daily new cases

```{r covid-daily-new-cases, echo=FALSE}

# Basic graph of new cases per day for the last x months

# Convert case data to a denormalised tibble. Pivot all columns except "Date" into a "country" column and a "cases" column. For PM data and all cases
covid_cases_p <- covid_cases %>% pivot_longer(!date, names_to = "country", values_to = "cases")

# Filter dataset for Malaysia only and for the last 6 months. Filter a bit longer for 7-day rolling average.
covid_cases_my <- filter(covid_cases_p, country == "Malaysia" & date > ymd("2020-10-15"))

# Add a 7-day average line to the data and filter again for the last 6 months
covid_cases_my <- covid_cases_my %>% mutate(cases_7ra = rollmean(cases, 7, fill = NA, align = "right")) %>% 
  filter(date > ymd("2020-10-31"))

# Plot plot plot
covid_cases_my_plot <- ggplot(data = covid_cases_my, aes(x = date)) +
  geom_col(aes(y = cases), color = "#FFDE00", fill = "#FFDE00") +
  geom_line(aes(y = cases_7ra), color = "#1252B3") +
  scale_x_date(name = "Date", date_breaks = "1 month", date_labels = "%b-%y") +
  scale_y_continuous(name = "Cases") +
  labs(title = "Daily new cases are trending up since late March", subtitle = "Daily new cases and 7-day rolling average of new cases", caption = "Source: Our World in Data.")

covid_cases_my_plot


```

## Vaccinations

Our World In Data uses the metric "Share of people who have received at least one dose of Covid-19 vaccine." Adding in "Share of population registered for vaccination." 

```{r covid-vaccinations, echo=FALSE}

# Load data (compiled by Faiz)
my_vaccinations <- read_csv("MY_Vaccination.csv", col_types = cols(
  Date = col_date(format = "%d-%m-%Y"), 
  category = col_character(), 
  Registrations = col_double(),
  First_dose = col_double(), 
  Second_dose = col_double(), 
  Total_doses = col_double()), skip_empty_rows = TRUE)

# Malaysian population
my_pop <- 32370000

# Pivot the data and remove lines where the date is NA
my_vaccinations <- my_vaccinations %>% pivot_longer(c("Registrations", "First_dose", "Second_dose", "Total_doses"), names_to = "category", values_to = "pop_count") %>% filter(!is.na(Date))

# Add a column with the percentage of population
my_vaccinations <- my_vaccinations %>% mutate(percentage = (pop_count / my_pop))

# Plot
my_vaccinations_p <- ggplot(data = filter(my_vaccinations, category %in% c("Registrations", "First_dose")), aes(x = Date)) +
  geom_line(aes(y = percentage, colour = category)) + 
  geom_hline(yintercept = 1, colour = "#999999") +
  scale_x_date(name = "Date", date_labels = "%d-%b-%y") +
  scale_y_continuous(name = "% of population", labels = scales::percent) + 
  scale_colour_brewer(palette = "Set1", labels = c("First dose", "Registrations")) +
  labs(title = "Vaccination registration is sluggish and vaccine roll-out starting slow", subtitle = "Percentage of population who registered for vaccine and received at least one dose", caption = "Source: Ministry of Health Malaysia") +
  theme(legend.position = "right", legend.title = element_blank())

my_vaccinations_p

```



 
## Tests, cases and positivity rate

Source: Our World In Data. 

```{r covid-19-linear, echo=FALSE}

# ===== Data wrangling =====

# Convert case data to a denormalised tibble. Pivot all columns except "Date" into a "country" column and a "cases" column. For PM data and all cases
covid_cases_p <- covid_cases %>% pivot_longer(!date, names_to = "country", values_to = "cases")
covid_cases_pivot <- covid_cases_pm %>% pivot_longer(!date, names_to = "country", values_to = "cases_pm")

# Filter COVID-19 data for Malaysia and South Korea, from 2020-10-31 to 2021-02-28

# Define filter values
iso_codes <- c("MYS", "KOR")
countries <- c("Malaysia", "South Korea")

date_from <- ymd("2020-09-30")
date_to <- ymd("2021-04-12")

# Define a tibble with the dates of MCO/CMCO restrictions
lockdowns <- tribble(
  ~lockdown, ~start, ~end,
  #--------|-------|-----
  "MCO-1", ymd("2020-03-18"), ymd("2020-05-04"),
  "CMCO-1", ymd("2020-05-04"), ymd("2020-06-09"),
  "CMCO-2", ymd("2020-10-14"), ymd("2020-12-05"),
  "MCO-2", ymd("2021-01-11"), ymd("2021-03-05"),
)

# Filter case data and rename columns for ease of use; add an ISO code column for easy matching
covid_cases_m <- filter(covid_cases_pivot, country %in% countries & date > date_from & date < date_to)
covid_cases_m <- rename(covid_cases_m, 
                        record.date = `date`,
                        daily.cases.per.million = `cases_pm`)
covid_cases_m <- covid_cases_m %>% mutate(ISO.code = case_when(
  country == "Malaysia" ~ "MYS",
  country == "South Korea" ~ "KOR",
  TRUE ~ country
))

# Filter tests data according to parameters and only useful columns; rename columns for ease of use
covid_tests_m <- filter(covid_tests, covid_tests$`ISO code` %in% iso_codes & covid_tests$Date > date_from & covid_tests$Date < date_to) %>% select(c("ISO code", "Date", "Daily change in cumulative total", "Daily change in cumulative total per thousand", "7-day smoothed daily change", "7-day smoothed daily change per thousand", "Short-term positive rate", "Short-term tests per case")) 
covid_tests_m <- rename(covid_tests_m, 
      ISO.code = `ISO code`, 
      record.date = `Date`,
      daily.tests = `Daily change in cumulative total`, 
      daily.tests.per.thousand = `Daily change in cumulative total per thousand`, 
      smoothed.daily.change = `7-day smoothed daily change`, 
      smoothed.daily.change.per.thousand = `7-day smoothed daily change per thousand`, 
      st.pos.rate = `Short-term positive rate`, 
      st.tests.per.case = `Short-term tests per case`)

# Filter NA's and sort rows in ascending order (by Date)
covid_cases_m <- arrange(covid_cases_m, record.date)
covid_tests_m <- arrange(covid_tests_m, record.date)

# Merge cases and tests data
covid_all <- covid_tests_m %>% left_join(covid_cases_m, by = c("ISO.code", "record.date"))


```

### Evolution of testing 

```{r covid-daily-tests}

# New facet labels for countries
country.labs <- c("South Korea", "Malaysia")
names(country.labs) <- c("KOR", "MYS")

# Plot daily tests (per thousand)
covid_dailytests_plot <- ggplot(data = covid_all, aes(x = record.date)) + 
  geom_bar(aes(y = daily.tests.per.thousand, colour = ISO.code), stat = "identity") + 
  scale_x_date(name = "Date") +
  scale_y_continuous(name = "Count") + 
  labs(title = "Daily tests per thousand for South Korea and Malaysia", caption = "From 01-Oct-2020 to 28-Feb-2021. Source: Our World In Data.") +
  theme(legend.position = "none")

covid_dailytests_plot + facet_wrap( ~ ISO.code)

```

### Evolution of daily cases

```{r covid-daily-cases}

# Plot daily cases (per million) 
covid_dailycases_plot <- ggplot(data = covid_all, aes(x = record.date)) + 
  geom_bar(aes(y = daily.cases.per.million, colour = ISO.code), stat = "identity") + 
  scale_x_date(name = "Date") +
  scale_y_continuous(name = "Count") + 
  labs(title = "Daily cases per million for South Korea and Malaysia", caption = "From 01-Oct-2020 to 28-Feb-2021. Source: Our World In Data.") +
  theme(legend.position = "none")

covid_dailycases_plot + facet_wrap( ~ ISO.code)

```

### Evolution of test positivity rate

This rate shows the short-term positivity rate of tests. 

```{r covid-posrate}

# Plot test positivity rate
covid_posrate_plot <- ggplot(data = covid_all) + 
  geom_line(aes(x = record.date, y = st.pos.rate, color = ISO.code)) +
  scale_x_date(name = "Date") +
  scale_y_continuous(labels = scales::percent, name = "Percent") + 
  labs(title = "Short-term test positivity rate for South Korea and Malaysia", caption = "From 01-Oct-2020 to 28-Feb-2021. Source: Our World In Data.") + 
  theme(legend.position = "none")

covid_posrate_plot + facet_wrap(~ ISO.code)

```

## Mobility

Analysis of mobility is based on Google's data for Malaysia. 
The charts below show the percent change compared to a baseline. The baseline is the median number of daily routing requests between early January to early February 2020. 

```{r mobility, echo=FALSE}

# Filter the raw data to only include nationwide change and correct date range
mobi_nation <- filter(mobi, is.na(sub_region_1) & date > date_from & date < date_to)
mobi_nation <- rename(mobi_nation, 
                      retail.rec = "retail_and_recreation_percent_change_from_baseline", 
                      grocery.pharmacy = "grocery_and_pharmacy_percent_change_from_baseline", 
                      parks = "parks_percent_change_from_baseline", 
                      transit = "transit_stations_percent_change_from_baseline", 
                      workplaces = "workplaces_percent_change_from_baseline", 
                      residential = "residential_percent_change_from_baseline")

# New facet labels for place categories
dest_cats.labs <- c("Retail & Rec", "Grocery & Pharmacy", "Parks", "Transit Stations", "Workplaces", "Residential")
names(dest_cats.labs) <- c("retail.rec", "grocery.pharmacy", "parks", "transit", "workplaces", "residential")

# The data is shown in tabular format, not very useful for analysis, so need to pivot the different categories into one column
mobi_nation <- mobi_nation %>% pivot_longer(c("retail.rec", "grocery.pharmacy", "parks", "transit", "workplaces", "residential"), names_to = "category", values_to = "change.from.baseline")

```

### Change from baseline across categories

This chart shows the change in destinations relative to a pre-pandemic baseline, smoothed using a 7-day rolling average. Each time movement restrictions are announced, there is a noticeable drop in retail and recreation destinations as well as transit stations. 

```{r mobi-plot-all, echo=FALSE, fig.asp=0.8}

# Filter the table to pick up only national data; Add a column with 7-day rolling average
mobi_nation <- mobi_nation %>% arrange(desc(category)) %>%
  group_by(category) %>%
  mutate(change.from.baseline.smooth = rollmean(change.from.baseline, 7, fill = NA)) %>%
  ungroup()

# plot the data; re-using the labels defined in the mobility block. 
mobi_plot <- ggplot(data = mobi_nation) +
  geom_rect(data = filter(lockdowns, start > date_from), inherit.aes = FALSE, aes(xmin = start, xmax = end, ymin = -Inf, ymax = +Inf), fill= 'red', alpha = 0.1) +
  geom_area(aes(x = date, y = change.from.baseline.smooth, color = category, fill = category), alpha = 0.33) +
  scale_x_date(name = "Date", date_label = "%b-%y") +
  scale_y_continuous(name = "Change from baseline (%)") +
  labs(title = "Routing requests to most destinations remain below baseline", subtitle = "Change in routing requests compared to baseline (Jan-2020 to Feb-2020)", caption = "Shaded areas indicate MCO/CMCO restrictions.\nSource: Google Covid-19 Community Mobility Reports.") +  
  theme(legend.position = "none")

mobi_plot + facet_wrap(~ category, labeller = labeller(category = dest_cats.labs))

```

The effect of the resumption of interstate travel on 2020-12-07 is well visible in the "Transit Stations" category. From that day there is a clear uptick in the number of requests for transit stations. Looking at that chart in isolation: 

```{r mobility-transit, echo=FALSE}

mobi_plot_transit <- ggplot(data = filter(mobi_nation, category == "transit")) +
  geom_area(aes(x = date, y = change.from.baseline), color = "#1252B3", fill = "#0D6EFF", alpha = 0.33) +
  scale_x_date(name = "Date") +
  scale_y_continuous(name = "Change from baseline (%)") +
  theme(legend.position = "none") +
  geom_vline(xintercept = ymd("2020-12-06")) +
  geom_text(mapping = aes(x = ymd("2020-12-06"), y = 0, label = "interstate travel", hjust = 0, vjust = 0)) +
  geom_vline(xintercept = ymd("2020-10-14")) +
  geom_text(mapping = aes(x = ymd("2020-10-14"), y = 0, label = "MCO 2.0", hjust = 0, vjust = 0)) + 
  geom_vline(xintercept = ymd("2021-01-13")) +
  geom_text(mapping = aes(x = ymd("2021-01-13"), y = 0, label = "MCO 3.0", hjust = 0, vjust = 0))

mobi_plot_transit

```

For further analysis, we could look to include economic data (high frequency indicators would be best), and relate that to the evolution of the pandemic and the mobility data. 

```{r covid-retail, echo=FALSE}

mobi_plot_transit <- ggplot(data = filter(mobi_nation, category %in% c("grocery.pharmacy", "retail.rec"))) +
  geom_rect(data = filter(lockdowns, start > date_from), inherit.aes = FALSE, aes(xmin = start, xmax = end, ymin = -Inf, ymax = +Inf), fill= 'red', alpha = 0.1) +
  geom_area(aes(x = date, y = change.from.baseline.smooth), color = "#1252B3", fill = "#1252B3", alpha = 0.5) +
  scale_x_date(name = "Date", date_labels = "%b-%y") +
  scale_y_continuous(name = "Change from baseline (%)") +
  scale_colour_brewer(palette = "Paired") +
  scale_fill_brewer(palette = "Paired") +
  labs(title = "Routing requests to Retail & Recreation destinations remain below baseline", subtitle = "Change in routing requests compared to baseline (Jan-2020 to Feb-2020)", caption = "Shaded areas indicate more stringent restrictions.\nSource: Google Covid-19 Community Mobility Reports.") +
  theme(legend.position = "none")

mobi_plot_transit + facet_wrap(~ category, labeller = labeller(category = dest_cats.labs))

```

## Passenger movements at MAHB airports

Passenger movements data provide an insight into the health of the tourism sector. Judging by the most recent data, travel levels are still far below the average levels of 2019. 

```{r mahb_ops, echo=FALSE}

# ===== MAHB Operating Statistics =====
# From MAHB website
mahb_ops <- read_csv("MAHBOperatingStatistics.csv")

# Define a tibble with the travel restriction dates
travel_res <- tribble(
  ~lockdown, ~start, ~end,
  #--------|-------|-----
  "MCO-1", ymd("2020-03-18"), ymd("2020-06-10"),
  "CMCO-1", ymd("2020-10-14"), ymd("2020-12-05"),
  "MCO-2", ymd("2021-01-11"), ymd("2021-03-31"),
)

mahb_ops_plot <- ggplot(data = mahb_ops, aes(x = mahb_ops$date)) +
  geom_rect(data = travel_res, inherit.aes = FALSE, aes(xmin = start, xmax = end, ymin=-Inf, ymax=+Inf), fill='red', alpha = 0.1) +
  geom_col(aes(y = mahb_ops$pax_mvmt_total, colour = airport, fill = airport)) +
  scale_x_date(name = "Date", date_breaks = "6 month", date_minor_breaks = "1 month", date_labels = "%b-%y") +
  scale_y_continuous(name = "Passenger movements (thousands)") + 
  scale_colour_brewer(palette = "Paired") +
  scale_fill_brewer(palette = "Paired") +
  labs(title = "MAHB airports total passenger movements are still far below 2019 levels", subtitle = "Total monthly passenger movement (thousands) From September 2019 to February 2021", caption = "Shaded areas indicate interstate travel restrictions in Klang Valley.\nSource: MAHB operating statistics.") +
  theme(legend.position = "right")

mahb_ops_plot

```
## Retail sales

Another leading indicator for economic performance, retail sales. 

```{r covid-retail-sales, echo=FALSE}

# Load data - from DOSM then turned into CSV data
retail_sales <- read_csv("MY_WholesaleRetail.csv")

# Pivot the sales data to longer form
retail_sales <- retail_sales %>% pivot_longer(c("Total", "Wholesale", "Retail", "Motor"), names_to = "category", values_to = "Sales")

# Filter for 18 months, same as passenger movement
retail_sales <- filter(retail_sales, Date > ymd("2020-01-31") & category %in% c("Wholesale", "Retail", "Motor"))

retail_sales_plot <- ggplot(data = retail_sales, aes(x = Date)) +
  geom_rect(data = lockdowns, inherit.aes = FALSE, aes(xmin = start, xmax = end, ymin=-Inf, ymax=+Inf), fill='red', alpha = 0.2) +
  geom_col(aes(y = Sales, colour = category, fill = category)) +
  scale_x_date(name = "Date", date_breaks = "6 month", date_minor_breaks = "1 month", date_labels = "%b-%y") +
  scale_y_continuous(name = "Retail sales (RM million)") + 
  scale_colour_brewer(palette = "Set1") +
  scale_fill_brewer(palette = "Set1") +
  labs(title = "Retail sales are weakening", subtitle = "Total monthly passenger movement (thousands) From September 2019 to February 2021", caption = "Shaded areas indicate interstate travel restrictions in Klang Valley.\nSource: MAHB operating statistics.") +
  theme(legend.position = "none")

retail_sales_plot + facet_wrap(~ category)

```

```{r covid-retail-sa, echo=FALSE}

# Load data - from DOSM then turned into CSV data
retail_sales_sa <- read_csv("MY_WholesaleRetail_SeasAdj.csv")

# Pivot the sales data to longer form
retail_sales_sa <- retail_sales_sa %>% pivot_longer(c("Total", "Wholesale", "Retail", "Motor"), names_to = "category", values_to = "Sales")

# Filter for 18 months, same as passenger movement
retail_sales_sa <- filter(retail_sales_sa, Date > ymd("2019-11-30") & category %in% c("Wholesale", "Retail", "Motor"))

# Build ancillary tibble for displaying 'high watermarks' on the facet charts
hline_max_sales <- tribble(
  ~category, ~max_sales,
  #--------|-------
  "Wholesale", filter(retail_sales_sa, category == "Wholesale") %>% arrange(desc(Sales)) %>% slice(1) %>% pull(Sales),
  "Retail", filter(retail_sales_sa, category == "Retail") %>% arrange(desc(Sales)) %>% slice(1) %>% pull(Sales),
  "Motor", filter(retail_sales_sa, category == "Motor") %>% arrange(desc(Sales)) %>% slice(1) %>% pull(Sales)
)

# Plot
retail_sales_sa_plot <- ggplot(data = retail_sales_sa, aes(x = Date)) +
  geom_rect(data = lockdowns, inherit.aes = FALSE, aes(xmin = start, xmax = end, ymin=-Inf, ymax=+Inf), fill='red', alpha = 0.1) +
  geom_line(aes(y = Sales, colour = category)) +
  geom_hline(yintercept = 100, colour = "#999999") +
  geom_hline(data = hline_max_sales, aes(yintercept = max_sales, colour = category), linetype = "dashed") +
  scale_x_date(name = "Date", date_labels = "%b-%y") +
  scale_y_continuous(name = "Volume Index") + 
  scale_colour_brewer(palette = "Set1") +
  scale_fill_brewer(palette = "Set1") +
  labs(title = "Retail sales have not yet recovered to previous levels", subtitle = "Seasonally adjusted volume index of wholesale and retail (2015=100)", caption = "Dashed line represents the 'high watermark' for each sector.\nShaded areas indicate MCO/CMCO restrictions.\nSource: DOSM.") +
  theme(legend.position = "none")

retail_sales_sa_plot + facet_wrap(~ category)


```

