---
title: "REFSA COVID-19 Malaysia Response"
author: "REFSA Research Team"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: 
  html_document: 
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

# Load libraries
library(tidyverse)
library(ggplot2)
library(scales)
library(lubridate)
library(dplyr)

# Load REFSA logo
# refsa_logo <- png::readPNG("Logo.png")

# Load data
# ===== Google Mobility Reports =====
# https://www.apple.com/covid19/mobility
# mobi <- read.csv("https://covid19-static.cdn-apple.com/covid19-mobility-data/2007HotfixDev48/v2/en-us/applemobilitytrends-2020-05-04.csv", na.strings = "", fileEncoding = "UTF-8-BOM")
mobi <- read_csv("2020_MY_Region_Mobility_Report.csv")

# ===== Covid test and cases (Our World in Data) =====
# https://www.ecdc.europa.eu/en/publications-data/download-todays-data-geographic-distribution-covid-19-cases-worldwide
covid_data <- read_csv("covid-19-daily-tests-vs-daily-new-confirmed-cases-per-million.csv")

```


## Objective

Malaysia is currently struggling to deal with the third wave of the Covid-19 pandemic. This analysis aims to highlight a few data points that illuminate the response so far of the government, and where additional investment may be required. 

 1. Test positivity rate, and the need to perform more testing
 2. Changes in mobility, related to the positivity rate
 
## Test positivity rate

The test positivity rate is still quite elevated in Malaysia, above the minimum threshold recommended by the WHO (5%), and especially high compared to a successful country in managing Covid-19, South Korea. The following charts depict the evolution of number of new cases, number of tests, and positivity rate. All data is 7-day smoothed (rolling average); people tested are per 1,000 population; new daily cases are per million population.

Source: Our World In Data. 

```{r covid-19-linear, echo=FALSE}

# Filter COVID-19 data for Malaysia and South Korea
geo_ids <- c("MYS", "KOR")
covid_data_m <- filter(covid_data, Code %in% geo_ids)

# No need to convert the date, R recognised it as a date field in the CSV
# Convert string date to Date
# covid_data_MYd <- covid_data_m %>% mutate(date = dmy(Date))

# Sort rows in ascending order (by Date)
covid_data_MYd <- arrange(covid_data_m, Date)

# Calculate cumulative number of cases and deaths - not required here, plot is relative
# covid_data_MYd <- covid_data_MYd %>% group_by(geoId) %>% mutate(cumCases = cumsum(cases))
# covid_data_MYd <- covid_data_MYd %>% group_by(geoId) %>% mutate(cumDeaths = cumsum(deaths))

# covid_data_clean <- na.omit(covid_data_MYd)

```

```{r covid-plot, echo=FALSE}

# Plot MY COVID-19 cases
covid_my_plot <- ggplot(data = covid_data_MYd) + 
  geom_line(aes(x = Date, y = DailyTests, color = Code, group = Code)) +
#  geom_line(aes(x = date, y = , color = geoId), linetype = "dashed") +
   scale_x_date(name = "Date")
covid_my_plot
```