---
title: "REFSA COVID-19 - Early 2022"
author: "REFSA Research Team"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: 
  html_document: 
    keep_md: true
fig.width: 6
fig.asp: 0.618
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

# Load libraries
library(tidyverse)
library(ggplot2)
library(scales)
library(lubridate)
library(dplyr)
library(zoo)
library(hrbrthemes)
library(RColorBrewer)

# Load REFSA logo
# refsa_logo <- png::readPNG("Logo.png")

# Load data

# Covid cases per million population
covid_cases_pm <- read_csv("https://raw.githubusercontent.com/owid/covid-19-data/master/public/data/jhu/new_cases_per_million.csv")

# ===== Covid test data (Our World in Data) =====
# https://github.com/owid/covid-19-data/blob/master/public/data/testing/covid-testing-all-observations.csv
covid_tests <- read_csv("https://raw.githubusercontent.com/owid/covid-19-data/master/public/data/testing/covid-testing-all-observations.csv")

```


This tracker brings together a number of oft-referred to statistics to monitor the evolution of the Covid-19 pandemic in Malaysia, and highlight some interesting data points. As of now, the following data is included: 

 1. Daily count of tests, cases and test positivity rate
 2. Mobility data, based on Google's Covid-19 location data
 
## Daily new cases

```{r covid-daily-new-cases, echo=FALSE}

# Basic graph of new cases per day for the last x months

# ===== Covid test and cases (Our World in Data) =====
# https://github.com/owid/covid-19-data/blob/master/public/data/jhu/new_cases_per_million.csv (Cases are per million population!)
# All Covid case data
covid_cases <- read_csv("https://raw.githubusercontent.com/owid/covid-19-data/master/public/data/jhu/new_cases.csv")

# Convert case data to a denormalised tibble. Pivot all columns except "Date" into a "country" column and a "cases" column. For PM data and all cases
covid_cases_p <- covid_cases %>% pivot_longer(!date, names_to = "country", values_to = "cases")

# Filter dataset for Malaysia only and for the last 6 months. Filter a bit longer for 7-day rolling average.
covid_cases_my <- filter(covid_cases_p, country == "Malaysia" & date > ymd("2021-08-01"))

# Add a 7-day average line to the data and filter again for the last 6 months
covid_cases_my <- covid_cases_my %>% mutate(cases_7ra = rollmean(cases, 7, fill = NA, align = "right")) %>% 
  filter(date > ymd("2021-08-01"))

# Plot plot plot
covid_cases_my_plot <- ggplot(data = covid_cases_my, aes(x = date)) +
  geom_col(aes(y = cases), colour = "#91B7D9", fill="#91B7D9") +
  geom_line(aes(y = cases_7ra), colour = "#2F5373") +
  scale_x_date(breaks = breaks_pretty(), labels = label_date_short()) +
  scale_y_continuous(name = "Cases") +
  labs(title = "New cases are increasing quickly", subtitle = "Daily new cases and 7-day rolling average of new cases", caption = "Source: Our World in Data.") +
  theme_minimal() + 
  theme(axis.title.x = element_blank())

covid_cases_my_plot


```

## Vaccinations

Vaccination data is based on Malaysia's MoH statistics. 

```{r covid-vax-data, echo=FALSE}

# Load vaccination data from MoH GitHub
my_vax <- read_csv("https://raw.githubusercontent.com/MoH-Malaysia/covid19-public/main/vaccination/vax_malaysia.csv")

# Load static data - population size
my_pop <- read_csv("https://raw.githubusercontent.com/MoH-Malaysia/covid19-public/main/static/population.csv")

```

```{r covid-vax-prep, echo = FALSE}

# Malaysian population pivot. Column 'pop' contains the total vaccinated population, this doesn't match the sum of the individual categories also included in the table. The individual categories 'pop_18' and 'pop_5' are for the respective vaccination groups. 
my_pop_all_vax <- my_pop %>% pivot_longer(c("pop", "pop_18", "pop_60", "pop_12", "pop_5"), names_to = "pop_group", values_to = "pop_count")

# Filter vax data for the columns we want
my_vax_all <- my_vax %>% select(date, cumul_partial, cumul_full, cumul_booster)

# Calculate net numbers for the chart areas to fit (otherwise we're double-counting across doses)
my_vax_all <- my_vax_all %>% mutate(net_first = cumul_partial - cumul_full) %>% mutate(net_second = cumul_full - cumul_booster)

# Pivot the data and remove lines where the date is NA
my_vaccinations <- my_vax_all %>% pivot_longer(c("cumul_partial", "cumul_full", "cumul_booster", "net_second", "net_first"), names_to = "category", values_to = "pop_count") %>% filter(!is.na(Date))

# Add a column with the percentage of population
my_pop_all = my_pop_all_vax %>% filter(pop_group == "pop" & state == "Malaysia") %>% pull(pop_count)
my_vaccinations <- my_vaccinations %>% mutate(percentage = (pop_count / my_pop_all))

```

```{r covid-vax-plot, echo=FALSE}

my_vax_data = filter(my_vaccinations, category %in% c("net_first", "net_second", "cumul_booster"))

# Plot
my_vaccinations_p <- ggplot(data = my_vax_data, aes(x = date)) +
#  geom_line(aes(y = percentage, colour = category)) +
  geom_area(aes(y = percentage, fill = factor(my_vax_data$category, levels=c("net_first", "net_second", "cumul_booster"), ordered = TRUE))) + 
  geom_hline(yintercept = 1, colour = "#999999") +
  scale_x_date(name = "Date", date_labels = "%b-%y") +
  scale_y_continuous(name = "% of population", labels = scales::percent) + 
  scale_fill_brewer(palette = "BuPu", labels = c("One dose", "Two doses", "Booster")) +
  labs(title = "Overall vaccination coverage is high but booster uptake sluggish", subtitle = "Percentage of population who received one dose, two doses, and booster", caption = "Source: Ministry of Health Malaysia") +
  theme_minimal() +
  theme(legend.position = "right", legend.title = element_blank(), axis.title.x = element_blank())

my_vaccinations_p

```

```{r covid-vax-group-data, echo = FALSE}

# Filter vax data for the columns we want
my_vax_group <- my_vax %>% select(date, cumul_partial, cumul_full, cumul_booster, cumul_partial_child, cumul_full_child, cumul_partial_adol, cumul_full_adol)

# Add a 'booster' column for each group, to have comparable metrics across all categories
my_vax_group <- my_vax_group %>%
  mutate(cumul_booster_adol = 0) %>%
  mutate(cumul_booster_child = 0) %>% 
  mutate(cumul_booster_adult = cumul_booster)

# Calculate net numbers for the chart areas to fit (otherwise we're double-counting across doses). 
# Compute a rate for adults as well, by subtracting the child and adol totals from the partial and full numbers for the entire adult group. 
my_vax_group <- my_vax_group %>% 
  mutate(net_first_adult = (cumul_partial - cumul_partial_adol - cumul_partial_child) - (cumul_full - cumul_full_adol - cumul_full_child)) %>%
  mutate(net_second_adult = (cumul_full - cumul_full_adol - cumul_full_child) - (cumul_booster - cumul_booster_adol - cumul_booster_child)) %>% 
  mutate(net_first_child = cumul_partial_child - cumul_full_child) %>% 
  mutate(net_second_child = cumul_full_child - cumul_booster_child) %>%
  mutate(net_first_adol = cumul_partial_adol - cumul_full_adol) %>%
  mutate(net_second_adol = cumul_full_adol - cumul_booster_adol)

# Pivot the data and remove lines where the date is NA. This gives us a table with one column 'category' and the net amounts for each. 
my_vax_group <- my_vax_group %>% pivot_longer(c("cumul_partial", "cumul_full", "cumul_booster", "net_second_adult", "net_first_adult", "cumul_booster_adult", "cumul_full_child", "cumul_partial_child", "cumul_full_adol", "cumul_partial_adol", "net_first_child", "net_second_child", "cumul_booster_child", "net_first_adol", "net_second_adol", "cumul_booster_adol"), names_to = "category", values_to = "pop_count") %>% filter(!is.na(Date))

# We need a second column with the group (adult, adol, child) to be able to facet the charts. Creating this column based on the category name
my_vax_group <- my_vax_group %>% mutate(group = case_when(
  str_detect(category, "adult") ~ "adult",
  str_detect(category, "adol") ~ "adol", 
  str_detect(category, "child") ~ "child",
  TRUE ~ "all"
  ))

# We need a third column for the metric, otherwise we can't facet the charts. For starters restricting to first and second dose, since children and adolescents do not yet have boosters. 
my_vax_group <- my_vax_group %>% mutate(dose = case_when(
  str_detect(category, "net_first") | str_detect(category, "partial") ~ "first",
  str_detect(category, "net_second") | str_detect(category, "full") ~ "second",
  str_detect(category, "booster") ~ "booster"
  ))

# Read population totals into variables for ease of use and speed
my_vax_pop_18 <- my_pop_all_vax %>% filter(pop_group == "pop_18" & state == "Malaysia") %>% pull(pop_count)
my_vax_pop_60 <- my_pop_all_vax %>% filter(pop_group == "pop_60" & state == "Malaysia") %>% pull(pop_count)
my_vax_pop_12 <- my_pop_all_vax %>% filter(pop_group == "pop_12" & state == "Malaysia") %>% pull(pop_count)
my_vax_pop_5 <- my_pop_all_vax %>% filter(pop_group == "pop_5" & state == "Malaysia") %>% pull(pop_count)

# Add a column with the percentage of population
my_vax_group <- my_vax_group %>% mutate(percentage = case_when(
  group == "adult" ~ pop_count / (my_vax_pop_18),  # According to docs, 60+ pop not broken out in vax stats
  group == "adol" ~ pop_count / (my_vax_pop_12),
  group == "child" ~ pop_count / (my_vax_pop_5)
  ))

```


```{r covid-vax-group-plot, echo = FALSE}

# Select only necessary rows (net rows, essentially) - also filter in time for last 3 months
# Find the latest date available in the data and go back 3 months
my_vax_max_date <- my_vax_group %>% summarise(date_to = max(date)) %>% pull(date_to)
date_start <- my_vax_max_date - months(3)
my_vax_group_data <- my_vax_group %>% filter(date > date_start)
my_vax_group_data <- my_vax_group_data %>% filter(category %in% c("net_first_adult", "net_second_adult", "cumul_booster_adult", "net_first_adol", "net_second_adol", "cumul_booster_adol", "net_first_child", "net_second_child", "cumul_booster_child"))

# Plot
my_vax_grp_p <- ggplot(data = my_vax_group_data, aes(x = date)) +
  geom_area(aes(y = percentage, fill = factor(dose, levels=c("first", "second", "booster"), ordered = TRUE))) + 
  geom_hline(yintercept = 1, colour = "#999999") +
  scale_x_date(name = "Date", labels = label_date_short(), breaks = breaks_pretty()) +
  scale_y_continuous(name = "% of population", labels = scales::percent) + 
  scale_fill_brewer(palette = "BuPu", labels = c("First", "Second", "Booster")) +
  labs(title = "Vaccination percentage in children and adolescents", subtitle = "Percentage of population who received one dose, two doses", caption = "Source: Ministry of Health Malaysia") +
  theme_minimal() +
  theme(legend.position = "right", legend.title = element_blank(), axis.title.x = element_blank()) +
  facet_wrap(~ factor(group, levels=c("adult", "adol", "child"), labels = c("18+", "12-17", "5-11"), ordered = TRUE))

my_vax_grp_p


```

## Mobility

Analysis of mobility is based on Google's data for Malaysia. 
The charts below show the percent change compared to a baseline. The baseline is the median number of daily routing requests between early January to early February 2020. 

```{r mobility, echo=FALSE}

# ===== Google Mobility Reports =====
# https://www.apple.com/covid19/mobility (alternative source)
# Google provides data in two files (one per year), so merging them together here too. 
mobi_20 <- read_csv("2020_MY_Region_Mobility_Report.csv")
mobi_21 <- read_csv("2021_MY_Region_Mobility_Report.csv")
mobi_22 <- read_csv("2022_MY_Region_Mobility_Report.csv")
mobi <- full_join(mobi_20, mobi_21)
mobi <- full_join(mobi, mobi_22)

# Filter the raw data to only include nationwide change and correct date range
# Hacked in hardcoded values for the dates; more elegant would be to make them dynamic, e.g. 6 month lookback window from the latest available date. 
mobi_nation <- filter(mobi, is.na(sub_region_1) & date > ymd("2021-09-01"))
mobi_nation <- rename(mobi_nation, 
                      retail.rec = "retail_and_recreation_percent_change_from_baseline", 
                      grocery.pharmacy = "grocery_and_pharmacy_percent_change_from_baseline", 
                      parks = "parks_percent_change_from_baseline", 
                      transit = "transit_stations_percent_change_from_baseline", 
                      workplaces = "workplaces_percent_change_from_baseline", 
                      residential = "residential_percent_change_from_baseline")

# New facet labels for place categories
dest_cats.labs <- c("Retail & Rec", "Grocery & Pharmacy", "Parks", "Transit Stations", "Workplaces", "Residential")
names(dest_cats.labs) <- c("retail.rec", "grocery.pharmacy", "parks", "transit", "workplaces", "residential")

# The data is shown in tabular format, not very useful for analysis, so need to pivot the different categories into one column
mobi_nation <- mobi_nation %>% pivot_longer(c("retail.rec", "grocery.pharmacy", "parks", "transit", "workplaces", "residential"), names_to = "category", values_to = "change.from.baseline")

```

### Change from baseline across categories

This chart shows the change in destinations relative to a pre-pandemic baseline, smoothed using a 7-day rolling average. Each time movement restrictions are announced, there is a noticeable drop in retail and recreation destinations as well as transit stations. 

```{r mobi-plot-all, echo=FALSE}

# Filter the table to pick up only national data; Add a column with 7-day rolling average
mobi_nation <- mobi_nation %>% arrange(desc(category)) %>%
  group_by(category) %>%
  mutate(change.from.baseline.smooth = rollmean(change.from.baseline, 7, fill = NA)) %>%
  ungroup()

# plot the data; re-using the labels defined in the mobility block. 
# Removed the display of lockdowns here since there were no changes in the period we looked at.
mobi_plot <- ggplot(data = mobi_nation) +
  geom_area(aes(x = date, y = change.from.baseline.smooth), colour = "#2F5373", fill = "#91B7D9", alpha = 0.33) +
  scale_x_date(name = "Date", breaks = breaks_pretty(), labels = label_date_short()) +
  scale_y_continuous(name = "Change from baseline (%)") +
  labs(title = "Routing requests to most destinations remain below baseline", subtitle = "Change in routing requests compared to baseline (Jan-2020 to Feb-2020)", caption = "Source: Google Covid-19 Community Mobility Reports.") +  
  theme_minimal() +
  theme(legend.position = "none", axis.title.x = element_blank()) +
  facet_wrap(~ category, labeller = labeller(category = dest_cats.labs)) +
  theme(panel.spacing.x = unit(4, "mm"))

mobi_plot 

```

For further analysis, we could look to include economic data (high frequency indicators would be best), and relate that to the evolution of the pandemic and the mobility data. 

## Passenger movements at MAHB airports

Passenger movements data provide an insight into the health of the tourism sector. Judging by the most recent data, travel levels are still far below the average levels of 2019. 

```{r mahb_ops, echo=FALSE}

# ===== MAHB Operating Statistics =====
# From MAHB website
mahb_ops <- read_delim("MAHBOperatingStatistics_2021.csv", delim = ";", locale = locale(decimal_mark = ",", grouping_mark = ".", date_format = "%d/%m/%Y"))

# Define a tibble with the travel restriction dates
travel_res <- tribble(
  ~lockdown, ~start, ~end,
  #--------|-------|-----
  "MCO-1", ymd("2020-03-18"), ymd("2020-06-10"),
  "CMCO-1", ymd("2020-10-14"), ymd("2020-12-05"),
  "MCO-2", ymd("2021-01-11"), ymd("2021-03-31"),
)

# Plot - showing KLIA and MASB passengers as categories, summing up to the total for the entire system in Malaysia.
# Also using a different palette that works better for the colours. 
mahb_ops_plot <- ggplot(data = mahb_ops, aes(x = date)) +
  geom_bar(aes(y = pax_mvmt_total, colour = airport, fill = airport), stat = "identity") +
  scale_x_date(breaks = breaks_pretty(), labels = label_date_short()) +
  scale_y_continuous(name = "Passenger movements (thousands)") + 
  scale_colour_brewer(palette = "Paired") +
  scale_fill_brewer(palette = "Paired") +
  theme_minimal() +
  labs(title = "Total passenger movements are still far below pre-pandemic levels", subtitle = "MAHB Airports total monthly passenger movement (thousands), September 2019 to December 2021", caption = "Source: MAHB operating statistics.") +
  theme(legend.position = "right", axis.title.x = element_blank(), legend.title = element_blank())

mahb_ops_plot

```   
## Retail sales

```{r covid-retail-sa, echo=FALSE}

# Load data - from DOSM then turned into CSV data
retail_sales_sa <- read_delim("MY_WholesaleRetail_SeasAdj_Jan20Dec21.csv", delim = ";", locale = locale(decimal_mark = ",", grouping_mark = ".", date_format = "%d/%m/%Y"))

# Pivot the sales data to longer form
retail_sales_sa <- retail_sales_sa %>% pivot_longer(c("Total", "Wholesale", "Retail", "Motor"), names_to = "category", values_to = "Sales")

# Filter for 18 months, same as passenger movement
retail_sales_sa <- filter(retail_sales_sa, category %in% c("Wholesale", "Retail", "Motor"))

# Build ancillary tibble for displaying 'high watermarks' on the facet charts
# Basically, we're looking for the max value in each category in the dataset we have, then extract it as the y value for the hline 
# corresponding to that section. The steps are basically to filter out each category, arrange from highest to lowest,
# extract the first record (highest value), then pull out the value
# (not using pull() would lead to a tibble within a tibble, which ggplot() can't deal with.)
hline_max_sales <- tribble(
  ~category, ~max_sales,
  #--------|-------
  "Wholesale", filter(retail_sales_sa, category == "Wholesale" & Date < ymd("2020-03-01")) %>% arrange(desc(Sales)) %>% slice(1) %>% pull(Sales),
  "Retail", filter(retail_sales_sa, category == "Retail" & Date < ymd("2020-03-01")) %>% arrange(desc(Sales)) %>% slice(1) %>% pull(Sales),
  "Motor", filter(retail_sales_sa, category == "Motor" & Date < ymd("2020-03-01")) %>% arrange(desc(Sales)) %>% slice(1) %>% pull(Sales)
)

# Define a tibble with the dates of MCO/CMCO restrictions
 lockdowns <- tribble(
   ~lockdown, ~start, ~end,
   #--------|-------|-----
   "MCO-1", ymd("2020-03-18"), ymd("2020-05-04"),
   "CMCO-1", ymd("2020-05-04"), ymd("2020-06-09"),
   "CMCO-2", ymd("2020-10-14"), ymd("2020-12-05"),
   "MCO-2", ymd("2021-01-11"), ymd("2021-03-05"),
 )

# Plot
retail_sales_sa_plot <- ggplot(data = retail_sales_sa %>% filter(Date > ymd("2019-12-01")), aes(x = Date)) +
  geom_rect(data = lockdowns, inherit.aes = FALSE, aes(xmin = start, xmax = end, ymin=-Inf, ymax=+Inf), fill='gray50', alpha = 0.1) +
  geom_line(aes(y = Sales), colour = "#2F5373") +
  geom_hline(yintercept = 100, colour = "#999999") +
  geom_hline(data = hline_max_sales, aes(yintercept = max_sales), colour = "#2F5373", linetype = "dashed") +
  scale_x_date(breaks = breaks_pretty(), labels = label_date_short()) +
  scale_y_continuous(name = "Volume Index") + 
  labs(title = "Retail sales show an uneven trajectory", subtitle = "Seasonally adjusted volume index of motor vehicle, wholesale and retail sales (2015=100)", caption = "Dashed line represents the 'high watermark' for each sector pre-pandemic.\nShaded areas indicate MCO/CMCO restrictions.\nSource: DOSM.") +
  theme_minimal() +
  theme(legend.position = "none", axis.title.x = element_blank()) +
  facet_wrap(~ factor(category, levels = c("Wholesale", "Retail", "Motor"), ordered = TRUE)) +
  theme(panel.spacing.x = unit(4, "mm"))
  
retail_sales_sa_plot


```


## School closures

Malaysia is among the countries where schools have been closed (fully or partially) the longest, when compared to ASEAN or other countries. The educational lag that the affected children would have incurred practically cannot be made up anymore, which bodes ill for their prospects. A shift in policy regarding schools is therefore urgently needed. 

```{r covid-school-closures, echo=FALSE}

# Load data; CSV file directly from UNESCO website
school_closure <- read_csv("https://en.unesco.org/sites/default/files/duration_school_closures.csv", skip=18)

# Rename columns to be more usable
school_closure <- rename(school_closure, 
              country = "Country",
              iso = "ISO",
              pandemic_total = "Duration of FULL and PARTIAL school closures (in weeks)",
              pandemic_full = "Duration of FULL closures (in weeks)",
              pandemic_partial = "Duration of PARTIAL closures (in weeks)",
              full_mar20_aug20 = "Duration of FULL closures from Mar-Aug 20 (in weeks)",
              full_sep20_aug21 = "Duration of FULL closures from Sep 20 - Aug 21 (in weeks)",
              full_sep21_oct21 = "Duration of FULL closures from Sep 21 - Oct 21 (in weeks)",
              partial_mar20_aug20 = "Duration of PARTIAL closures from Mar-Aug 20 (in weeks)",
              partial_sep20_aug21 = "Duration of PARTIAL closures from Sep 20 - Aug 21 (in weeks)",
              partial_sep21_oct21 = "Duration of PARTIAL closures from Sep 21 - Oct 21 (in weeks)")

# Define ASEAN countries for comparison
asean_iso_codes <- c("BRN", "KHM", "IDN", "LAO", "MYS", "MMR", "PHL", "SGP", "THA", "VNM")

# Indicate which countries are in ASEAN
school_closure <- school_closure %>% mutate(region = ifelse(iso %in% asean_iso_codes, "ASEAN", NA))

# Calculate average closure durations for ASEAN
sc_asean_avg <- school_closure %>% filter(region == "ASEAN") %>% summarise(country = "ASEAN", iso = "ASEAN", region = "ASEAN", across(pandemic_total:partial_sep21_oct21, mean))

# Calculate average closure durations for the world
sc_world_avg <- school_closure %>% summarise(country = "World", iso = "World", region = "World", across(pandemic_total:partial_sep21_oct21, mean))

# Bind rows with averages to the country data
sc <- bind_rows(school_closure, sc_asean_avg, sc_world_avg)

# Pivot table for easier graphing
sc <- sc %>% pivot_longer(c("pandemic_total", "pandemic_full", "pandemic_partial", "full_mar20_aug20", "full_sep20_aug21", "full_sep21_oct21", "partial_mar20_aug20", "partial_sep20_aug21", "partial_sep21_oct21"), names_to = "metric", values_to = "duration")

# Categorise closure duration in type: partial or full
sc <- sc %>% mutate(type = case_when(
  str_detect(metric, "partial") ~ "partial",
  str_detect(metric, "full") ~ "full", 
  str_detect(metric, "total") ~ "total"))

# Categorise closure period
sc <- sc %>% mutate(period = case_when(
  str_detect(metric, "mar20_aug20") ~ "mar20-aug20",
  str_detect(metric, "sep20_aug21") ~ "sep20-aug21",
  str_detect(metric, "sep21_oct21") ~ "sep21-oct21",
  str_detect(metric, "pandemic") ~ "pandemic",
))

# Plot
sc_p <- ggplot(data = sc %>% filter(iso %in% c("MYS", "ASEAN", "World", "SGP") & type %in% c("full", "partial") & period == "pandemic"), aes(x = factor(iso, levels=c("MYS", "ASEAN", "World", "SGP"), labels = c("Malaysia", "ASEAN", "World", "Singapore"), ordered = TRUE))) +
  geom_col(aes(y = duration, fill = factor(type, levels=c("partial", "full"), labels = c("Partial closure", "Full closure"), ordered = TRUE))) + 
  scale_y_continuous(name = "Closure duration (weeks)") + 
  scale_fill_brewer(palette = "Paired") +
  labs(title = "School closures in Malaysia are longer than ASEAN and world averages", subtitle = "Duration in weeks of full and partial school closures", caption = "Source: UNESCO") + 
  theme_minimal() +
  theme(legend.position = "right", legend.title = element_blank(), axis.title.x = element_blank())

sc_p

```
